Getting and Cleaning Data Course Project
==============================================

Daniel Bell
----------------------------------------------

### Samsung Galaxy S acceleramotor data - Codebook


#### Variables contained in the tidy dataset
Name             | Description
-----------------|------------
subject          | The ID of the subject who was studied
activity         | name of the activity the subject is currently performing
featDomain       | Boolean indicating indicating either time or frequency
featInstrument   | Instrument used to measure signal
featAcceleration | Acceleration signal
featVariable     | Either the mean or standard deviation of observation
featJerk         | Jerk signal
featMagnitude    | Magnitude of the signals
featAxis         | 3-axial signals in the X, Y and Z directions (X, Y, or Z)
featCount        | Count of data points used to compute average
featAverage      | Average of each variable for each activity and each subject

#### Summary of Variables
```{r, echo=TRUE}
summary(dataTidy)
```

#### A sample of the Data
```{r, echo=TRUE}
head(dataTidy)
```

#### Structure of the Tidy Dataset
```{r, echo=TRUE}
str(dataTidy)
```

#### Keys
```{r, echo=TRUE}
key(dataTidy)
```

#### Output Dataset
The tidy dataset was created in the run_analysis.R script and is saved in the same location as the script.


 
